---
title: "Midterm Project: DATA474 Qian"
author: "Weslee Nguyen and Sedona O'Hearn"
format: pdf
editor: visual
---

# Libraries

```{r}
#| echo: false
#| warning: false

library(tidyverse)
library(dplyr)
library(ggplot2)
library(mice)     # For imputation
library(igraph)   # To later visualize highly correlated vars
library(corrplot) # Correlation heatmap
library(purrr)    # Like mapply, lapply, sapply
library(glmnet)   # Fits LASSO for logistic regression
library(caret)    # Confusion matrix
library(MASS)     # Stepwise Subset Selection
library(pROC)     # Evaluate models using ROC curves and AUC
library(nnet)
```

# CHF Classification

## Step #1: Preprocessing

```{r}
#| eval: FALSE

albumin_creatinine <- read.csv("Datasets/albumin-creatinine.csv")
biochem_panel <- read.csv("Datasets/biochem_panel.csv")
blood_pressure <- read.csv("Datasets/blood_pressure.csv")
body_measures <- read.csv("Datasets/body_measures.csv")
CBC <- read.csv("Datasets/CBC.csv")
cigarettes <- read.csv("Datasets/cigarettes.csv")
demographics <- read.csv("Datasets/demographics.csv")
diabetes <- read.csv("Datasets/diabetes.csv")
diet_second_day <- read.csv("Datasets/diet_second_day.csv")
environmental_pollutants <- read.csv("Datasets/environmental_pollutants.csv")
medical_conditions <- read.csv("Datasets/medical_conditions.csv")
mental_health <- read.csv("Datasets/mental_health.csv")
physical_activity <- read.csv("Datasets/physical_activity.csv")

datasets <- list(
  albumin_creatinine,
  biochem_panel,
  blood_pressure,
  body_measures,
  CBC,
  cigarettes,
  demographics,
  diabetes,
  diet_second_day,
  environmental_pollutants,
  medical_conditions,
  mental_health,
  physical_activity
)

# There's column X in all of them, get rid of it
datasets_clean <- map(datasets, ~ select(.x, -any_of("X")))

# Merge all datasets by unique participant identifier SEQN
NHANES <- reduce(datasets_clean, inner_join, by = "SEQN")

# Only retain adults and people who answered Y/N to CHF
NHANES <- NHANES %>% filter(RIDAGEYR >= 18, MCQ160B %in% c(1, 2)) %>%
  mutate(CHF = if_else(MCQ160B == 1, 1, 0)) %>%
  select(-MCQ160B)

# Factorize for classification; 1 = Has CHF, 0 = No CHF
NHANES$CHF <- factor(NHANES$CHF, levels = c(0, 1))

# Drop empty predictors (>30% NAs)
NA_vars <- NHANES %>%
  summarise(across(everything(), ~ mean(is.na(.), na.rm = FALSE))) %>%
  pivot_longer(
    cols = everything(), 
    names_to = "variable", 
    values_to = "NA_percent") %>%
  arrange(desc(NA_percent))

# Many variables tossed as they're kid-specific e.g. head circumference
# BMI Variables to Toss: BMXRECUM, BMIRECUM, BMXHEAD, BMDBMIC, BMIHT, BMIARM, 
      # BMIARMC, BMIWAIST, BMIWT, BMIHIP, BMILEG
# Smoking Variables: SMQ040, SMQ621, SMD630, SMD641, SMD650, SMD100MN
# Demographics: RIDAGEMN, RIDEXAGM, RIDEXPRG, DMDYRUSR, DMDHRGND, DMDHRAGZ
      # DMDHREDZ, DMDHRMAZ, DMDHSEDZ
# Medical Conditions: MCQ149 (Liver),  MCQ500 (Cancer), MCQ230A-D (Cancer)
      # MCQ510A-F (Cancer, all), MCQ035 + MCQ040 (Asthma), MCQ050 (Thyroid)
      # MCQ170L/M (Liver + Thyroid), MCQ195 (Arthritis)
# Diabetes: DIQ040 (Age of Dx), DIQ050 (insulin), DIQ060 + DIQ060U (Meds), DIQ070
# Mental Health: DPQ100 (Depression score)
# Physical Activity: PAD820 (Leisurely Activity)
# Diet Recall: DR2SKY (Sugar intake), DR2DBIH (Dietary intake)
# Lab work: LBXSLDSI (Serum Sodium)

vars_to_keep <- NA_vars %>%
  filter(NA_percent <= 0.3) %>%
  pull(variable)

NHANES <- NHANES %>% select(any_of(vars_to_keep))

##### MULTICOLLINEARITY: Remove vars basically the same, else MICE crashes
numeric_vars <- NHANES %>% select(where(is.numeric))
cor_matrix <- cor(numeric_vars, use = "pairwise.complete.obs") # Avoid NAs

cor_pairs <- cor_matrix %>%
  as.data.frame() %>%
  rownames_to_column(var = "var1") %>%
  pivot_longer(-var1, names_to = "var2", values_to = "cor") %>%
  filter(var1 != var2) %>%          # remove diagonal as cor = 1
  filter(abs(cor) > 0.95) %>%       # threshold cor = 0.95
  arrange(desc(abs(cor)))

cor_pairs

# Turn cor_pairs into graph to see connected components
# Remove from each component till there's only one member
edges <- cor_pairs %>% select(var1, var2)
g <- graph_from_data_frame(edges, directed = FALSE)
clusters <- components(g)
clusters$membership

# There's 30 clusters = remove at least 30 variables
# Simply chose 1 per cluster
# Any lab variables = starts with LBX; diet = DR2, body measure = LBD, urine = URX
NHANES <- NHANES %>% select(-c(LBXSTB, LBDSGBSI, LBDSTPSI, LBDSALSI, URXUMS, 
                 URXCRS, WTPH2YR.y, WTPH2YR.x, LBDBMNSI, LBDBCDSI, LBDSCRSI, 
                 LBDTHGSI, LBDSUASI, LBDSCHSI, LBDSTRSI, LBDBPBSI, LBDBSESI, 
                 LBDSBUSI, LBDSPHSI, LBDSIRSI, DR2TM181, DR2TP182, DR2DRSTZ, 
                 LBDSGLSI, DR2TP226, DR2TS160, DR2TFDFE, LBXHCT, DR2TP225, 
                 DR2TS060, DR2TS140, LBDSCASI, RIDRETH3, BPXOPLS1, WTMEC2YR, 
                 DR2TTFAT, DR2TS180, PAD790U, PAD810U, BPAOARM, DRABF, 
                 LBDBPBLC, LBDBSELC, LBDBMNLC, URDUCRLC))

##### IMPUTATION: Use MICE to imputate remaining missing data
# Handles continuous and categorical data
# Treat NA variable as dependent, regress on all other variables
# Predictive Mean Matching (PMM): Find other cases w/ similar predicted values, 
      # then randomly choose one of their actual observed values for imputation--
      # thus assumes data is missing at random (no pattern)
# m = 5 imputed datasets made for stabilized guessing

imputation <- mice(NHANES, m = 1, method = "pmm", seed = 123)
NHANES <- complete(imputation)

# Export to make sure it's never ruined!
NHANES_cleaned <- write.csv(NHANES, "NHANES_cleaned.csv")
```

Documentation:

-   [Albumin-Creatinine](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/ALB_CR_L.htm)

-   [Biochemistry Panel](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/BIOPRO_L.htm)

-   [Blood Pressure](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/BPXO_L.htm)

-   [Body Measures](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/BMX_L.htm)

-   [Complete Blood Count](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/CBC_L.htm)

-   [Cigarettes](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/SMQ_L.htm)

-   [Demographics](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DEMO_L.htm#SEQN)

-   [Diabetes](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DIQ_L.htm)

-   [Diet](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DR2TOT_L.htm#DR2TP182)

-   [Environmental Pollutants](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/PBCD_L.htm)

-   [Medical Conditions](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/MCQ_L.htm#MCQ160b)

-   [Mental Health](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DPQ_L.htm#DPQ070)

-   [Physical Activity](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/PAQ_L.htm)

## Step #2: Exploratory Data Analysis

```{r}

NHANES_cleaned <- read.csv("NHANES_cleaned.csv")

################################################################################
################################ FIGURE I ######################################
################################################################################

# I) Make mini-correlation heatmap related to CHF

cor_with_chf <- cor(NHANES_cleaned)["CHF", ]
cor_with_chf <- sort(abs(cor_with_chf), decreasing = TRUE)
top10_predictors <- names(cor_with_chf)[2:11]
top10_predictors

# MCQ160E: Heart Attack
# MCQ160P: COPD/Emphysema/Chronic Bronchitis
# DIQ010: Diabetes
# RIDAGEYR: Age
# BMXWAIST: Waist Circumference
# LBXSCR: Serum Creatinine
# LBXRDW: Red Blood Cell Distribution Width
# LBXSBU: Blood Urea Nitrogen
# LBXSCH: Serum Cholesterol
# LBXSUA: Serum Uric Acid

mini_NHANES <- NHANES_cleaned %>%
  dplyr::select(MCQ160E, MCQ160P, DIQ010, RIDAGEYR, BMXWAIST,
         LBXSCR, LBXRDW, LBXSBU, LBXSCH, LBXSUA, CHF) %>%
  mutate(
    MCQ160E = ifelse(MCQ160E == 1, 1, 0),
    MCQ160P = ifelse(MCQ160P == 1, 1, 0),
    DIQ010 = ifelse(DIQ010 == 1, 1, 0)
  ) %>%
  rename(
    Heart_Attack = MCQ160E,
    Chronic_Breathing_Issues = MCQ160P,
    Diabetes = DIQ010,
    Age = RIDAGEYR,
    Waist_Circumference = BMXWAIST,
    Serum_Creatinine = LBXSCR,
    RBC_Distribution_Width = LBXRDW,
    Blood_Urea_Nitrogen = LBXSBU,
    Serum_Cholesterol = LBXSCH,
    Serum_Uric_Acid = LBXSUA
  )

pretty_labels <- c(
  "CHF",
  "Heart Attack",
  "Chronic Breathing Issues",
  "Diabetes",
  "Age (years)",
  "Waist Circumference (cm)",
  "Serum Creatinine (mg/dL)",
  "RBC Distribution Width (%)",
  "BUN (mg/dL)",
  "Serum Cholesterol (mg/dL)",
  "Serum Uric Acid (mg/dL)"
)

cor_matrix <- cor(mini_NHANES)
rownames(cor_matrix) <- pretty_labels
colnames(cor_matrix) <- pretty_labels
corrplot(cor_matrix, method = "color", type = "upper", addCoef.col = "black", 
         tl.srt = 45, tl.cex = 0.8, number.cex = 0.5)

################################################################################
################################################################################
################################################################################

# II) Histograms of Variables Most Correlated with CHF

pretty_labels <- c(
  Heart_Attack = "Heart Attack",
  Chronic_Breathing_Issues = "Chronic Breathing Issues",
  Diabetes = "Diabetes",
  Age = "Age (yrs)",
  Waist_Circumference = "Waist Circumference (cm)",
  Serum_Creatinine = "Serum Creatinine (mg/dL)",
  RBC_Distribution_Width = "RBC Distribution Width (%)",
  Blood_Urea_Nitrogen = "Blood Urea Nitrogen (mg/dL)",
  Serum_Cholesterol = "Serum Cholesterol (mg/dL)",
  Serum_Uric_Acid = "Serum Uric Acid (mg/dL)"
)

NHANES_pivot <- mini_NHANES %>% pivot_longer(
  cols = everything(), names_to = "variable", values_to = "value"
)

ggplot(NHANES_pivot, aes(x = value)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  facet_wrap(~variable, scales = "free", nrow = 3,
             labeller = labeller(variable = pretty_labels)) +
  labs(
    y = "Count",
    title = "Histogram of Variables Most Correlated with CHF"
  ) +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        strip.text = element_text(size = 6),
        plot.title = element_text(hjust = 0.5))

################################################################################
################################################################################
################################################################################

# III) Look at the 3 binary variables given CHF = 1 v. 0
binary_vars <- mini_NHANES %>% dplyr::select(Heart_Attack, Chronic_Breathing_Issues, 
                                      Diabetes, CHF)

long_binary <- binary_vars %>%
  pivot_longer(
    cols = c(Heart_Attack, Chronic_Breathing_Issues, Diabetes),
    names_to = "variable",
    values_to = "value"
  )

plot_data <- long_binary %>%
  group_by(variable, value) %>%
  summarise(
    CHF_rate = mean(CHF, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

ggplot(plot_data, aes(x = factor(value), y = CHF_rate, fill = factor(value))) +
  geom_col(linewidth = 1, color = "black") +
  facet_wrap(~variable, scales = "free_x", 
             labeller = labeller(variable = pretty_labels)) +
  scale_fill_manual(
    values = c("0" = "skyblue", "1" = "salmon"),
    labels = c("0" = "Never Had/Have (0)", "1" = "Had/Have (1)")
  ) +
  labs(
    y = "Proportion",
    fill = "CHF",
    title = "Proportion of Medical Issues Having Had v. Not Had CHF"
  ) +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.position = "bottom")

################################################################################
################################################################################
################################################################################

# IV) Look at continuous variables for CHF = 1 v. 0

continuous_vars <- mini_NHANES %>% dplyr::select(CHF, Age, Waist_Circumference, 
         Serum_Creatinine, RBC_Distribution_Width, Blood_Urea_Nitrogen, 
         Serum_Cholesterol, Serum_Uric_Acid)

long_cont <- continuous_vars %>%
  pivot_longer(
    cols = -CHF,
    names_to = "variable",
    values_to = "value"
  )

ggplot(long_cont, aes(x = factor(CHF), y = value, fill = factor(CHF))) +
  geom_boxplot(linewidth = 0.3, color = "black") +
  facet_wrap(~variable, scales = "free_y", nrow = 3,
             labeller = labeller(variable = pretty_labels)) +
  scale_fill_manual(
    values = c("0" = "skyblue", "1" = "salmon"),
    labels = c("0" = "Never Had/Have (0)", "1" = "Had/Have (1)")
  ) +
  labs(
    y = "Value",
    fill = "CHF",
    title = "Correlated Continuous Measures Had v. Not Had CHF"
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5),
  )
```

That gets four subfigures for the main EDA figure.

```{r}

################################################################################
################################ TABLE I #######################################
################################################################################

# Mean / Proportion for CHF = 0, 1 per variable

vars <- colnames(mini_NHANES)
vars <- vars[vars != "CHF"]

summarize_var <- function(data, var) {
  
  df <- data %>% dplyr::select(CHF, all_of(vars))
  
  CHF0 <- df %>% filter(CHF == 0) %>% pull(var)
  CHF1 <- df %>% filter(CHF == 1) %>% pull(var)
  
  mean0 <- mean(CHF0)
  mean1 <- mean(CHF1)
  
  se0 <- sd(CHF0) / sqrt(length(CHF0))
  se1 <- sd(CHF1) / sqrt(length(CHF1))
  
  test <- t.test(CHF0, CHF1)

  tibble(
    Variable = var,
    Mean0 = mean0,
    SE0 = se0,
    Mean1 = mean1,
    SE1 = se1,
    Ratio = mean1 / mean0,
    t_stat = test$statistic,
    p_value = test$p.value,
    Mean_diff = mean1 - mean0,
    CI_lower = test$conf.int[1],
    CI_upper = test$conf.int[2]
  )
}

summary_table <- map_dfr(vars, ~ summarize_var(mini_NHANES, .x))
summary_table
```

That gets the necessary information to populate Table I.

## Splitting the Data

I used createDataPartition from Caret instead of initial_split() from tidymodels (which I noticed was used for the tree later) because my computer was being evil. But when we combine this all into 1, we can pick one or the other method and it shouldn't really change anything.

```{r}
train_index <- createDataPartition(
  NHANES_cleaned$CHF,
  p = 0.80,
  list = FALSE
)

train_data <- NHANES_cleaned[train_index, ]

test_data  <- NHANES_cleaned[-train_index, ]
```

## Binary Logistic Regression -\> Neural Network

The imbalanced ratio of CHF cases to non CHF cases makes it highly difficult to classify accurately. Because of this, many models tend to favor predicting the majority non-CHF class, resulting in high overall accuracy but poor sensitivity and recall for true CHF cases. To address this issue, we first used logistic regression as a feature selection and screening method to identify the most clinically relevant predictors associated with CHF. These predictors were then used as inputs into a neural network model.

### Create Logistic Regression Model

```{r}
run_logistic_model <- function(train_data, test_data) {

  # Fit logistic regression
  logistic_model <- glm(
    CHF ~ .,
    data = train_data,
    family = binomial
  )

  # Predicted probabilities
  pred <- predict(
    logistic_model,
    newdata = test_data,
    type = "response"
  )

  # Convert probabilities into classes
  pred_class <- ifelse(pred > 0.5, 1, 0)

  pred_class <- factor(pred_class, levels = c(0,1))
  y_factor <- factor(test_data$CHF, levels = c(0,1))

  cm <- confusionMatrix(
    pred_class,
    y_factor,
    positive = "1"
  )

  list(
    model = logistic_model,
    confusion = cm,
    pred_prob = pred,
    pred_class = pred_class
  )
}
```

### Select Predictors

Variables for the neural network were selected using the binary logistic regression model. Predictors with the strongest statistical significance (p \< 0.001) were retained as inputs to the neural network in order to reduce dimensionality while preserving the variables most associated with CHF classification.

```{r}
logistic_full <- run_logistic_model(NHANES_cleaned, NHANES_cleaned)

coef_df <- summary(logistic_full$model)$coefficients %>%
  as.data.frame()

coef_df$Variable <- rownames(coef_df)

coef_df <- coef_df %>%
  filter(Variable != "(Intercept)") %>%
  arrange(`Pr(>|z|)`)

top_coef_df <- coef_df %>%
  slice_head(n = 15)

top_coef_df$log_p <- -log10(top_coef_df$`Pr(>|z|)`)

coef_plot <- ggplot(
  top_coef_df,
  aes(
    x = reorder(Variable, log_p),
    y = log_p
  )
) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Most Statistically Significant Logistic Regression Predictors",
    x = "Predictor",
    y = "-log10(p-value)"
  ) +
  theme_minimal()

ggsave(
  "logistic_feature_importance.png",
  plot = coef_plot,
  width = 8,
  height = 6,
  dpi = 300
)
```

```{r}
selected_vars <- c(

  "DR2TVB6",

  "DR2TMAGN",

  "DR2TWSZ",

  "LBXSCH",

  "LBXRDW",

  "MCQ160E",

  "RIDAGEYR",

  "SDMVSTRA"

)

```

### Create Neural Network

While logistic regression provides strong interpretability, it assumes a linear relationship between predictors and the log-odds of CHF. Neural networks allow more flexible nonlinear decision boundaries and may improve predictive performance by capturing more complex relationships among predictors.

```{r}
run_neural_network_model <- function(train_data, test_data, selected_vars,
                                     size = 5, decay = 0.01, threshold = 0.5) {
  
  X_train <- train_data[, selected_vars]
  X_test  <- test_data[, selected_vars]
  
  preproc <- preProcess(X_train, method = c("center", "scale"))
  
  X_train_scaled <- predict(preproc, X_train)
  X_test_scaled  <- predict(preproc, X_test)
  
  train_scaled <- data.frame(X_train_scaled, CHF = train_data$CHF)
  test_scaled  <- data.frame(X_test_scaled, CHF = test_data$CHF)
  
  set.seed(123)
  
  nn_model <- nnet(
    CHF ~ .,
    data = train_scaled,
    size = size,
    decay = decay,
    maxit = 500,
    trace = FALSE
  )
  
  pred <- predict(nn_model, newdata = test_scaled, type = "raw")
  pred <- as.numeric(pred)
  
  pred_class <- ifelse(pred > threshold, 1, 0)
  pred_class <- factor(pred_class, levels = c(0, 1))
  
  y_factor <- factor(test_scaled$CHF, levels = c(0, 1))
  
  cm <- confusionMatrix(pred_class, y_factor, positive = "1")
  
  auc_value <- as.numeric(auc(roc(test_scaled$CHF, pred)))
  
  list(
    model = nn_model,
    selected_vars = selected_vars,
    size = size,
    decay = decay,
    threshold = threshold,
    confusion = cm,
    pred_prob = pred,
    pred_class = pred_class,
    AUC = auc_value
  )
}
```

```{r}
tune_neural_network <- function(train_data, test_data, selected_vars) {
  sizes <- c(3, 5, 8)
  decays <- c(0.001, 0.01, 0.1)
  thresholds <- c(0.3, 0.4, 0.5)

  results <- data.frame()
  for (s in sizes) {
    for (d in decays) {
      for (t in thresholds) {
        nn <- run_neural_network_model(
          train_data,
          test_data,
          selected_vars,
          size = s,
          decay = d,
          threshold = t
        )
        cm <- nn$confusion
        results <- rbind(
          results,
          data.frame(
            Size = s,
            Decay = d,
            Threshold = t,
            Accuracy = as.numeric(cm$overall["Accuracy"]),
            Precision = as.numeric(cm$byClass["Pos Pred Value"]),
            Recall = as.numeric(cm$byClass["Sensitivity"]),
            F1 = as.numeric(cm$byClass["F1"]),
            AUC = nn$AUC
          )
        )
      }
    }
  }
  results
}
```

The neural network hyperparameters were tuned over a small search space to improve classification performance on the imbalanced CHF dataset. Hidden layer size, weight decay regularization, and classification threshold were systematically varied to evaluate their impact on model accuracy, precision, recall, and AUC. Because identifying true CHF cases was prioritized over maximizing overall accuracy, the final model was selected using the F1 score, which balances precision and recall. Lower classification thresholds generally improved sensitivity and recall by increasing the number of predicted CHF cases, though at the cost of additional false positives.

```{r}
nn_tuning_results <- tune_neural_network(
  train_data,
  test_data,
  selected_vars
)

nn_tuning_results$F1 <- with(
  nn_tuning_results,
  2 * Precision * Recall / (Precision + Recall)
)

best_params <- nn_tuning_results[which.max(nn_tuning_results$F1), ]

best_params
```

```{r}
###############################
# Final Neural Network Model
###############################

final_nn <- run_neural_network_model(
  train_data,
  test_data,
  selected_vars,
  size = best_params$Size,
  decay = best_params$Decay,
  threshold = best_params$Threshold
)

###############################
# Final Evaluation
###############################

final_nn$confusion
confusionMatrix(
  final_nn$pred_class,
  factor(test_data$CHF, levels = c(0,1)),
  positive = "1"
)

cm_table <- as.data.frame(final_nn$confusion$table)
conf_matrix <- ggplot(cm_table, aes(x = Reference, y = Prediction, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Freq), size = 6) +
  labs(
    title = "Test Confusion Matrix for Neural Network",
    x = "True CHF Status",
    y = "Predicted CHF Status"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggsave(
  "nn_confusion_matrix.png",
  plot = conf_matrix,
  width = 8,
  height = 6,
  dpi = 300
)
```

```{r}
final_nn$confusion$byClass
```

## 
