---
title: "Midterm Project: DATA474 Qian"
author: "Weslee Nguyen and Sedona O'Hearn"
format: pdf
editor: visual
---

# Libraries

```{r}
#| echo: false
#| warning: false

library(tidyverse)
library(dplyr)
library(ggplot2)
library(mice)     # For imputation
library(igraph)   # To later visualize highly correlated vars
library(corrplot) # Correlation heatmap
library(purrr)    # Like mapply, lapply, sapply
library(glmnet)   # Fits LASSO for logistic regression
library(caret)    # Confusion matrix
library(MASS)     # Stepwise Subset Selection
library(pROC)     # Evaluate models using ROC curves and AUC
```

# CHF Classification

## Step #1: Preprocessing

```{r}
#| eval: FALSE

albumin_creatinine <- read.csv("Datasets/albumin-creatinine.csv")
biochem_panel <- read.csv("Datasets/biochem_panel.csv")
blood_pressure <- read.csv("Datasets/blood_pressure.csv")
body_measures <- read.csv("Datasets/body_measures.csv")
CBC <- read.csv("Datasets/CBC.csv")
cigarettes <- read.csv("Datasets/cigarettes.csv")
demographics <- read.csv("Datasets/demographics.csv")
diabetes <- read.csv("Datasets/diabetes.csv")
diet_second_day <- read.csv("Datasets/diet_second_day.csv")
environmental_pollutants <- read.csv("Datasets/environmental_pollutants.csv")
medical_conditions <- read.csv("Datasets/medical_conditions.csv")
mental_health <- read.csv("Datasets/mental_health.csv")
physical_activity <- read.csv("Datasets/physical_activity.csv")

datasets <- list(
  albumin_creatinine,
  biochem_panel,
  blood_pressure,
  body_measures,
  CBC,
  cigarettes,
  demographics,
  diabetes,
  diet_second_day,
  environmental_pollutants,
  medical_conditions,
  mental_health,
  physical_activity
)

# There's column X in all of them, get rid of it
datasets_clean <- map(datasets, ~ select(.x, -any_of("X")))

# Merge all datasets by unique participant identifier SEQN
NHANES <- reduce(datasets_clean, inner_join, by = "SEQN")

# Only retain adults and people who answered Y/N to CHF
NHANES <- NHANES %>% filter(RIDAGEYR >= 18, MCQ160B %in% c(1, 2)) %>%
  mutate(CHF = if_else(MCQ160B == 1, 1, 0)) %>%
  select(-MCQ160B)

# Factorize for classification; 1 = Has CHF, 0 = No CHF
NHANES$CHF <- factor(NHANES$CHF, levels = c(0, 1))

# Drop empty predictors (>30% NAs)
NA_vars <- NHANES %>%
  summarise(across(everything(), ~ mean(is.na(.), na.rm = FALSE))) %>%
  pivot_longer(
    cols = everything(), 
    names_to = "variable", 
    values_to = "NA_percent") %>%
  arrange(desc(NA_percent))

# Many variables tossed as they're kid-specific e.g. head circumference
# BMI Variables to Toss: BMXRECUM, BMIRECUM, BMXHEAD, BMDBMIC, BMIHT, BMIARM, 
      # BMIARMC, BMIWAIST, BMIWT, BMIHIP, BMILEG
# Smoking Variables: SMQ040, SMQ621, SMD630, SMD641, SMD650, SMD100MN
# Demographics: RIDAGEMN, RIDEXAGM, RIDEXPRG, DMDYRUSR, DMDHRGND, DMDHRAGZ
      # DMDHREDZ, DMDHRMAZ, DMDHSEDZ
# Medical Conditions: MCQ149 (Liver),  MCQ500 (Cancer), MCQ230A-D (Cancer)
      # MCQ510A-F (Cancer, all), MCQ035 + MCQ040 (Asthma), MCQ050 (Thyroid)
      # MCQ170L/M (Liver + Thyroid), MCQ195 (Arthritis)
# Diabetes: DIQ040 (Age of Dx), DIQ050 (insulin), DIQ060 + DIQ060U (Meds), DIQ070
# Mental Health: DPQ100 (Depression score)
# Physical Activity: PAD820 (Leisurely Activity)
# Diet Recall: DR2SKY (Sugar intake), DR2DBIH (Dietary intake)
# Lab work: LBXSLDSI (Serum Sodium)

vars_to_keep <- NA_vars %>%
  filter(NA_percent <= 0.3) %>%
  pull(variable)

NHANES <- NHANES %>% select(any_of(vars_to_keep))

##### MULTICOLLINEARITY: Remove vars basically the same, else MICE crashes
numeric_vars <- NHANES %>% select(where(is.numeric))
cor_matrix <- cor(numeric_vars, use = "pairwise.complete.obs") # Avoid NAs

cor_pairs <- cor_matrix %>%
  as.data.frame() %>%
  rownames_to_column(var = "var1") %>%
  pivot_longer(-var1, names_to = "var2", values_to = "cor") %>%
  filter(var1 != var2) %>%          # remove diagonal as cor = 1
  filter(abs(cor) > 0.95) %>%       # threshold cor = 0.95
  arrange(desc(abs(cor)))

cor_pairs

# Turn cor_pairs into graph to see connected components
# Remove from each component till there's only one member
edges <- cor_pairs %>% select(var1, var2)
g <- graph_from_data_frame(edges, directed = FALSE)
clusters <- components(g)
clusters$membership

# There's 30 clusters = remove at least 30 variables
# Simply chose 1 per cluster
# Any lab variables = starts with LBX; diet = DR2, body measure = LBD, urine = URX
NHANES <- NHANES %>% select(-c(LBXSTB, LBDSGBSI, LBDSTPSI, LBDSALSI, URXUMS, 
                 URXCRS, WTPH2YR.y, WTPH2YR.x, LBDBMNSI, LBDBCDSI, LBDSCRSI, 
                 LBDTHGSI, LBDSUASI, LBDSCHSI, LBDSTRSI, LBDBPBSI, LBDBSESI, 
                 LBDSBUSI, LBDSPHSI, LBDSIRSI, DR2TM181, DR2TP182, DR2DRSTZ, 
                 LBDSGLSI, DR2TP226, DR2TS160, DR2TFDFE, LBXHCT, DR2TP225, 
                 DR2TS060, DR2TS140, LBDSCASI, RIDRETH3, BPXOPLS1, WTMEC2YR, 
                 DR2TTFAT, DR2TS180, PAD790U, PAD810U, BPAOARM, DRABF, 
                 LBDBPBLC, LBDBSELC, LBDBMNLC, URDUCRLC))

##### IMPUTATION: Use MICE to imputate remaining missing data
# Handles continuous and categorical data
# Treat NA variable as dependent, regress on all other variables
# Predictive Mean Matching (PMM): Find other cases w/ similar predicted values, 
      # then randomly choose one of their actual observed values for imputation--
      # thus assumes data is missing at random (no pattern)
# m = 5 imputed datasets made for stabilized guessing

imputation <- mice(NHANES, m = 1, method = "pmm", seed = 123)
NHANES <- complete(imputation)

# Export to make sure it's never ruined!
NHANES_cleaned <- write.csv(NHANES, "NHANES_cleaned.csv")
```

Documentation:

-   [Albumin-Creatinine](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/ALB_CR_L.htm)

-   [Biochemistry Panel](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/BIOPRO_L.htm)

-   [Blood Pressure](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/BPXO_L.htm)

-   [Body Measures](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/BMX_L.htm)

-   [Complete Blood Count](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/CBC_L.htm)

-   [Cigarettes](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/SMQ_L.htm)

-   [Demographics](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DEMO_L.htm#SEQN)

-   [Diabetes](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DIQ_L.htm)

-   [Diet](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DR2TOT_L.htm#DR2TP182)

-   [Environmental Pollutants](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/PBCD_L.htm)

-   [Medical Conditions](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/MCQ_L.htm#MCQ160b)

-   [Mental Health](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DPQ_L.htm#DPQ070)

-   [Physical Activity](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/PAQ_L.htm)

## Step #2: Exploratory Data Analysis

```{r}

NHANES_cleaned <- read.csv("NHANES_cleaned.csv")

################################################################################
################################ FIGURE I ######################################
################################################################################

# I) Make mini-correlation heatmap related to CHF

cor_with_chf <- cor(NHANES_cleaned)["CHF", ]
cor_with_chf <- sort(abs(cor_with_chf), decreasing = TRUE)
top10_predictors <- names(cor_with_chf)[2:11]
top10_predictors

# MCQ160E: Heart Attack
# MCQ160P: COPD/Emphysema/Chronic Bronchitis
# DIQ010: Diabetes
# RIDAGEYR: Age
# BMXWAIST: Waist Circumference
# LBXSCR: Serum Creatinine
# LBXRDW: Red Blood Cell Distribution Width
# LBXSBU: Blood Urea Nitrogen
# LBXSCH: Serum Cholesterol
# LBXSUA: Serum Uric Acid

mini_NHANES <- NHANES_cleaned %>%
  dplyr::select(MCQ160E, MCQ160P, DIQ010, RIDAGEYR, BMXWAIST,
         LBXSCR, LBXRDW, LBXSBU, LBXSCH, LBXSUA, CHF) %>%
  mutate(
    MCQ160E = ifelse(MCQ160E == 1, 1, 0),
    MCQ160P = ifelse(MCQ160P == 1, 1, 0),
    DIQ010 = ifelse(DIQ010 == 1, 1, 0)
  ) %>%
  rename(
    Heart_Attack = MCQ160E,
    Chronic_Breathing_Issues = MCQ160P,
    Diabetes = DIQ010,
    Age = RIDAGEYR,
    Waist_Circumference = BMXWAIST,
    Serum_Creatinine = LBXSCR,
    RBC_Distribution_Width = LBXRDW,
    Blood_Urea_Nitrogen = LBXSBU,
    Serum_Cholesterol = LBXSCH,
    Serum_Uric_Acid = LBXSUA
  )

pretty_labels <- c(
  "CHF",
  "Heart Attack",
  "Chronic Breathing Issues",
  "Diabetes",
  "Age (years)",
  "Waist Circumference (cm)",
  "Serum Creatinine (mg/dL)",
  "RBC Distribution Width (%)",
  "BUN (mg/dL)",
  "Serum Cholesterol (mg/dL)",
  "Serum Uric Acid (mg/dL)"
)

cor_matrix <- cor(mini_NHANES)
rownames(cor_matrix) <- pretty_labels
colnames(cor_matrix) <- pretty_labels
corrplot(cor_matrix, method = "color", type = "upper", addCoef.col = "black", 
         tl.srt = 45, tl.cex = 0.8, number.cex = 0.5)

################################################################################
################################################################################
################################################################################

# II) Histograms of Variables Most Correlated with CHF

pretty_labels <- c(
  Heart_Attack = "Heart Attack",
  Chronic_Breathing_Issues = "Chronic Breathing Issues",
  Diabetes = "Diabetes",
  Age = "Age (yrs)",
  Waist_Circumference = "Waist Circumference (cm)",
  Serum_Creatinine = "Serum Creatinine (mg/dL)",
  RBC_Distribution_Width = "RBC Distribution Width (%)",
  Blood_Urea_Nitrogen = "Blood Urea Nitrogen (mg/dL)",
  Serum_Cholesterol = "Serum Cholesterol (mg/dL)",
  Serum_Uric_Acid = "Serum Uric Acid (mg/dL)"
)

NHANES_pivot <- mini_NHANES %>% pivot_longer(
  cols = everything(), names_to = "variable", values_to = "value"
)

ggplot(NHANES_pivot, aes(x = value)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  facet_wrap(~variable, scales = "free", nrow = 3,
             labeller = labeller(variable = pretty_labels)) +
  labs(
    y = "Count",
    title = "Histogram of Variables Most Correlated with CHF"
  ) +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        strip.text = element_text(size = 6),
        plot.title = element_text(hjust = 0.5))

################################################################################
################################################################################
################################################################################

# III) Look at the 3 binary variables given CHF = 1 v. 0
binary_vars <- mini_NHANES %>% dplyr::select(Heart_Attack, Chronic_Breathing_Issues, 
                                      Diabetes, CHF)

long_binary <- binary_vars %>%
  pivot_longer(
    cols = c(Heart_Attack, Chronic_Breathing_Issues, Diabetes),
    names_to = "variable",
    values_to = "value"
  )

plot_data <- long_binary %>%
  group_by(variable, value) %>%
  summarise(
    CHF_rate = mean(CHF, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

ggplot(plot_data, aes(x = factor(value), y = CHF_rate, fill = factor(value))) +
  geom_col(linewidth = 1, color = "black") +
  facet_wrap(~variable, scales = "free_x", 
             labeller = labeller(variable = pretty_labels)) +
  scale_fill_manual(
    values = c("0" = "skyblue", "1" = "salmon"),
    labels = c("0" = "Never Had/Have (0)", "1" = "Had/Have (1)")
  ) +
  labs(
    y = "Proportion",
    fill = "CHF",
    title = "Proportion of Medical Issues Having Had v. Not Had CHF"
  ) +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.position = "bottom")

################################################################################
################################################################################
################################################################################

# IV) Look at continuous variables for CHF = 1 v. 0

continuous_vars <- mini_NHANES %>% dplyr::select(CHF, Age, Waist_Circumference, 
         Serum_Creatinine, RBC_Distribution_Width, Blood_Urea_Nitrogen, 
         Serum_Cholesterol, Serum_Uric_Acid)

long_cont <- continuous_vars %>%
  pivot_longer(
    cols = -CHF,
    names_to = "variable",
    values_to = "value"
  )

ggplot(long_cont, aes(x = factor(CHF), y = value, fill = factor(CHF))) +
  geom_boxplot(linewidth = 0.3, color = "black") +
  facet_wrap(~variable, scales = "free_y", nrow = 3,
             labeller = labeller(variable = pretty_labels)) +
  scale_fill_manual(
    values = c("0" = "skyblue", "1" = "salmon"),
    labels = c("0" = "Never Had/Have (0)", "1" = "Had/Have (1)")
  ) +
  labs(
    y = "Value",
    fill = "CHF",
    title = "Correlated Continuous Measures Had v. Not Had CHF"
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5),
  )
```

That gets four subfigures for the main EDA figure.

```{r}

################################################################################
################################ TABLE I #######################################
################################################################################

# Mean / Proportion for CHF = 0, 1 per variable

vars <- colnames(mini_NHANES)
vars <- vars[vars != "CHF"]

summarize_var <- function(data, var) {
  
  df <- data %>% dplyr::select(CHF, all_of(vars))
  
  CHF0 <- df %>% filter(CHF == 0) %>% pull(var)
  CHF1 <- df %>% filter(CHF == 1) %>% pull(var)
  
  mean0 <- mean(CHF0)
  mean1 <- mean(CHF1)
  
  se0 <- sd(CHF0) / sqrt(length(CHF0))
  se1 <- sd(CHF1) / sqrt(length(CHF1))
  
  test <- t.test(CHF0, CHF1)

  tibble(
    Variable = var,
    Mean0 = mean0,
    SE0 = se0,
    Mean1 = mean1,
    SE1 = se1,
    Ratio = mean1 / mean0,
    t_stat = test$statistic,
    p_value = test$p.value,
    Mean_diff = mean1 - mean0,
    CI_lower = test$conf.int[1],
    CI_upper = test$conf.int[2]
  )
}

summary_table <- map_dfr(vars, ~ summarize_var(mini_NHANES, .x))
summary_table
```

That gets the necessary information to populate Table I.

## Step #3: Screen-and-Stepwise Model

Due to nearly having 200 variables, one can significantly reduce the number of predictors before testing. One can employ PCA or some other dimensionality reduction, however that means the new predictors (being components) and subsequent units becomes uninterpretable. Alternatively, one can use the odds ratio. While it doesn't test significance, does keep predictors where the population of those with CHF has a different enough magnitude from those without. That way, the number of predictors is significantly reduced to test run a subset selection. Those best selection is absurd with $2^{27}$ viable models, one that merges front and backward stepwise selection serves as a balance to choose a model towards a relatively small AIC, a con being that insignificant predictors may still be included.

```{r}
run_stepwise_model <- function(train_data, test_data) {
  
  X_names <- setdiff(names(train_data), "CHF")
  Y <- train_data$CHF
  
  # Get rid of weak odds ratio
  screen_OR <- sapply(X_names, function(var) {
    formula <- as.formula(paste("CHF ~", var))
    model <- glm(formula, data = train_data, family = binomial)
    exp(coef(model)[2])  # OR
  })
  
  # Get if odds ratio > 1.5
  selected_vars <- names(screen_OR[abs(screen_OR - 1) > 0.5])
  selected_vars <- unique(selected_vars)
  
  # OR duplicates e.g. OSQ230 -> OSQ230.OSQ230, so trim
  selected_vars <- sub("\\..*$", "", selected_vars)
  selected_vars <- selected_vars[!is.na(selected_vars)]
  
  # Safety fallback
  if (length(selected_vars) == 0) {
    selected_vars <- X_names
  }
  
  # Initialize full model
  full_model <- as.formula(
    paste("CHF ~", paste(selected_vars, collapse = " + "))
  )
  
  final_data <- train_data[, c(selected_vars, "CHF")]
  initial_model <- glm(full_model, data = train_data, family = binomial)
  
  # AIC Stepwise Selection
  final_model <- stepAIC(initial_model, direction = "both", trace = FALSE)
  
  pred <- predict(final_model, newdata = test_data, type = "response")
  pred_class <- ifelse(pred > 0.5, 1, 0)
  pred_class <- factor(pred_class, levels = c(0,1))
  y_factor <- factor(test_data$CHF, levels = c(0,1))
  
  cm <- confusionMatrix(pred_class, y_factor, positive = "1")
  
  list(
    model = final_model,
    selected_vars = names(coef(final_model))[-1],
    confusion = cm,
    pred_prob = pred,
    pred_class = pred_class
  )
}

step_full <- run_stepwise_model(NHANES_cleaned, NHANES_cleaned)

summary(step_full$model)
length(step_full$selected_vars)
step_full$confusion
```

## Step #4: LASSO Logistic Regression

Alternatively, one can use all (nearly) 200 predictors, however that is not only overwhelming, but risks lack of practical interpretability (with too many factors to consider for CHF). Thus, logistic regression would benefit using LASSO to aggressively enforce parsimony.

```{r}
run_lasso_model <- function(train_data, test_data) {
  
  # Initialize matrices
  X_train <- as.matrix(train_data[, colnames(train_data) != "CHF"])
  Y_train <- train_data$CHF
  
  X_test <- as.matrix(test_data[, colnames(test_data) != "CHF"])
  Y_test <- test_data$CHF
  
  # Meant to show how coefficients die as lambda increases
  LASSO_model <- glmnet(X_train, Y_train, family = "binomial", alpha = 1)
  
  # Do 10-fold CV fitting bunch of LASSO Logistics
  set.seed(123)
  kfold_CV <- cv.glmnet(X_train, Y_train, family = "binomial", alpha = 1, nfolds = 10)
  
  # Predict from best model
  pred <- predict(kfold_CV, newx = X_test, s = "lambda.min", type = "response")
  pred <- as.numeric(pred)
  
  pred_class <- ifelse(pred > 0.5, 1, 0)
  pred_class <- factor(pred_class, levels = c(0,1))
  y_factor <- factor(Y_test, levels = c(0,1))
  
  cm <- confusionMatrix(pred_class, y_factor, positive = "1")
  
  list(
    model = LASSO_model,
    cv_model = kfold_CV,
    lambda_min = kfold_CV$lambda.min,
    lambda_1se = kfold_CV$lambda.1se,
    coefficients = coef(kfold_CV, s = "lambda.min"),
    confusion = cm,
    pred_prob = pred,
    pred_class = pred_class
  )
}

lasso_full <- run_lasso_model(NHANES_cleaned, NHANES_cleaned)

lasso_full$lambda.min
lasso_full$lambda.1se
lasso_full$coefficients
lasso_full$confusion
```

## Step #5: Model Evaluation

Model performance was evaluated using 5-fold cross-validation, where both models were trained on a subset of the data and tested on held-out folds. Accuracy, precision, recall, and AUC were computed for each fold and averaged to assess performance. Given the class imbalance in CHF prevalence, particular emphasis was placed on recall and AUC, as accuracy alone may be misleading.

```{r}
set.seed(123)

# Make 5 folds
folds <- createFolds(factor(NHANES_cleaned$CHF), k = 5, list = TRUE)

# Store results
step_metrics <- data.frame(
  Accuracy = numeric(5),
  Precision = numeric(5),
  Recall = numeric(5),
  AUC = numeric(5)
)

lasso_metrics <- data.frame(
  Accuracy = numeric(5),
  Precision = numeric(5),
  Recall = numeric(5),
  AUC = numeric(5)
)

# Optional: store full model objects
step_results <- vector("list", 5)
lasso_results <- vector("list", 5)

for (i in 1:5) {
  
  cat("Running fold", i, "\n")
  
  test_idx <- folds[[i]]
  train_data <- NHANES_cleaned[-test_idx, ]
  test_data  <- NHANES_cleaned[test_idx, ]
  
  # Run stepwise
  step_results[[i]] <- run_stepwise_model(train_data, test_data)
  
  # Run lasso
  lasso_results[[i]] <- run_lasso_model(train_data, test_data)
  
  # Stepwise metrics
  step_metrics$Accuracy[i] <- as.numeric(step_results[[i]]$confusion$overall["Accuracy"])
  step_metrics$Precision[i] <- as.numeric(step_results[[i]]$confusion$byClass["Pos Pred Value"])
  step_metrics$Recall[i] <- as.numeric(step_results[[i]]$confusion$byClass["Sensitivity"])
  step_metrics$AUC[i] <- as.numeric(
    auc(roc(test_data$CHF, step_results[[i]]$pred_prob))
  )
  
  # LASSO metrics
  lasso_metrics$Accuracy[i] <- as.numeric(lasso_results[[i]]$confusion$overall["Accuracy"])
  lasso_metrics$Precision[i] <- as.numeric(lasso_results[[i]]$confusion$byClass["Pos Pred Value"])
  lasso_metrics$Recall[i] <- as.numeric(lasso_results[[i]]$confusion$byClass["Sensitivity"])
  lasso_metrics$AUC[i] <- as.numeric(
    auc(roc(test_data$CHF, lasso_results[[i]]$pred_prob))
  )
}
```

```{r}
step_cv_results <- colMeans(step_metrics, na.rm = TRUE)
lasso_cv_results <- colMeans(lasso_metrics, na.rm = TRUE)

step_cv_results
lasso_cv_results
```

```{r}
cv_results <- rbind(
  Stepwise = step_cv_results,
  LASSO = lasso_cv_results
)

round(cv_results, 4)
```

## Step #6: ROC Curves

```{r}
# Get predictions from last fold
roc_step <- roc(test_data$CHF, step_results[[5]]$pred_prob)
roc_lasso <- roc(test_data$CHF, lasso_results[[5]]$pred_prob)

plot(roc_step, col = "blue", main = "ROC Curve Comparison")
lines(roc_lasso, col = "red")

legend("bottomright",
       legend = c("Stepwise", "LASSO"),
       col = c("blue", "red"),
       lwd = 2)
```
