---
instal---
title: "Midterm Project: DATA474 Qian"
author: "Weslee Nguyen, Sedona O'Hearn, and Jacob Stevens"
format: pdf
editor: visual
---

# Libraries

library(rpart.plot)

library(rpart)

```{r}
#| echo: false
#| warning: false

# For Step #1: Preprocessing, I don't use since the dataset is already made
library(mice)       # For imputation
library(igraph)     # To later visualize highly correlated vars
library(corrplot)   # Correlation heatmap

# For Step #2 onwards!
library(tidyverse)
library(caret)      # Confusion matrix
library(pROC)       # Evaluate models using ROC curves and AUC
library(vip)        # Variable Importance Plots (VIP)
library(e1071)      # For SVM
library(kernlab)    # For SVM as well

library(rpart.plot)
library(rpart)      # Decision Trees (first one I got when looking it up for R)
  # https://www.geeksforgeeks.org/r-language/decision-tree-in-r-programming/
  # Followed this tutorial - Weslee

library(tidymodels) # Preprocessing pipeline for modeling
  # Already has libraries purrr, dplyr, ggplot2, recipes, tidyr
  # Will have to update a lot of libraries to use this

library(vip)        # Quantify importance per feature
```

# CHF Classification

## Step #1: Preprocessing

```{r}
#| eval: FALSE

albumin_creatinine <- read.csv("Datasets/albumin-creatinine.csv")
biochem_panel <- read.csv("Datasets/biochem_panel.csv")
blood_pressure <- read.csv("Datasets/blood_pressure.csv")
body_measures <- read.csv("Datasets/body_measures.csv")
CBC <- read.csv("Datasets/CBC.csv")
cigarettes <- read.csv("Datasets/cigarettes.csv")
demographics <- read.csv("Datasets/demographics.csv")
diabetes <- read.csv("Datasets/diabetes.csv")
diet_second_day <- read.csv("Datasets/diet_second_day.csv")
environmental_pollutants <- read.csv("Datasets/environmental_pollutants.csv")
medical_conditions <- read.csv("Datasets/medical_conditions.csv")
mental_health <- read.csv("Datasets/mental_health.csv")
physical_activity <- read.csv("Datasets/physical_activity.csv")

datasets <- list(
  albumin_creatinine,
  biochem_panel,
  blood_pressure,
  body_measures,
  CBC,
  cigarettes,
  demographics,
  diabetes,
  diet_second_day,
  environmental_pollutants,
  medical_conditions,
  mental_health,
  physical_activity
)

# There's column X in all of them, get rid of it
datasets_clean <- map(datasets, ~ select(.x, -any_of("X")))

# Merge all datasets by unique participant identifier SEQN
NHANES <- reduce(datasets_clean, inner_join, by = "SEQN")

# Only retain adults and people who answered Y/N to CHF
NHANES <- NHANES %>% filter(RIDAGEYR >= 18, MCQ160B %in% c(1, 2)) %>%
  mutate(CHF = if_else(MCQ160B == 1, 1, 0)) %>%
  select(-MCQ160B)

# Factorize for classification; 1 = Has CHF, 0 = No CHF
NHANES$CHF <- factor(NHANES$CHF, levels = c(0, 1))

# Drop empty predictors (>30% NAs)
NA_vars <- NHANES %>%
  summarise(across(everything(), ~ mean(is.na(.), na.rm = FALSE))) %>%
  pivot_longer(
    cols = everything(), 
    names_to = "variable", 
    values_to = "NA_percent") %>%
  arrange(desc(NA_percent))

# Many variables tossed as they're kid-specific e.g. head circumference
# BMI Variables to Toss: BMXRECUM, BMIRECUM, BMXHEAD, BMDBMIC, BMIHT, BMIARM, 
      # BMIARMC, BMIWAIST, BMIWT, BMIHIP, BMILEG
# Smoking Variables: SMQ040, SMQ621, SMD630, SMD641, SMD650, SMD100MN
# Demographics: RIDAGEMN, RIDEXAGM, RIDEXPRG, DMDYRUSR, DMDHRGND, DMDHRAGZ
      # DMDHREDZ, DMDHRMAZ, DMDHSEDZ
# Medical Conditions: MCQ149 (Liver),  MCQ500 (Cancer), MCQ230A-D (Cancer)
      # MCQ510A-F (Cancer, all), MCQ035 + MCQ040 (Asthma), MCQ050 (Thyroid)
      # MCQ170L/M (Liver + Thyroid), MCQ195 (Arthritis)
# Diabetes: DIQ040 (Age of Dx), DIQ050 (insulin), DIQ060 + DIQ060U (Meds), DIQ070
# Mental Health: DPQ100 (Depression score)
# Physical Activity: PAD820 (Leisurely Activity)
# Diet Recall: DR2SKY (Sugar intake), DR2DBIH (Dietary intake)
# Lab work: LBXSLDSI (Serum Sodium)

vars_to_keep <- NA_vars %>%
  filter(NA_percent <= 0.3) %>%
  pull(variable)

NHANES <- NHANES %>% select(any_of(vars_to_keep))

##### MULTICOLLINEARITY: Remove vars basically the same, else MICE crashes
numeric_vars <- NHANES %>% select(where(is.numeric))
cor_matrix <- cor(numeric_vars, use = "pairwise.complete.obs") # Avoid NAs

cor_pairs <- cor_matrix %>%
  as.data.frame() %>%
  rownames_to_column(var = "var1") %>%
  pivot_longer(-var1, names_to = "var2", values_to = "cor") %>%
  filter(var1 != var2) %>%          # remove diagonal as cor = 1
  filter(abs(cor) > 0.95) %>%       # threshold cor = 0.95
  arrange(desc(abs(cor)))

cor_pairs

# Turn cor_pairs into graph to see connected components
# Remove from each component till there's only one member
edges <- cor_pairs %>% select(var1, var2)
g <- graph_from_data_frame(edges, directed = FALSE)
clusters <- components(g)
clusters$membership

# There's 30 clusters = remove at least 30 variables
# Simply chose 1 per cluster
# Any lab variables = starts with LBX; diet = DR2, body measure = LBD, urine = URX
NHANES <- NHANES %>% select(-c(LBXSTB, LBDSGBSI, LBDSTPSI, LBDSALSI, URXUMS, 
                 URXCRS, WTPH2YR.y, WTPH2YR.x, LBDBMNSI, LBDBCDSI, LBDSCRSI, 
                 LBDTHGSI, LBDSUASI, LBDSCHSI, LBDSTRSI, LBDBPBSI, LBDBSESI, 
                 LBDSBUSI, LBDSPHSI, LBDSIRSI, DR2TM181, DR2TP182, DR2DRSTZ, 
                 LBDSGLSI, DR2TP226, DR2TS160, DR2TFDFE, LBXHCT, DR2TP225, 
                 DR2TS060, DR2TS140, LBDSCASI, RIDRETH3, BPXOPLS1, WTMEC2YR, 
                 DR2TTFAT, DR2TS180, PAD790U, PAD810U, BPAOARM, DRABF, 
                 LBDBPBLC, LBDBSELC, LBDBMNLC, URDUCRLC))

##### IMPUTATION: Use MICE to imputate remaining missing data
# Handles continuous and categorical data
# Treat NA variable as dependent, regress on all other variables
# Predictive Mean Matching (PMM): Find other cases w/ similar predicted values, 
      # then randomly choose one of their actual observed values for imputation--
      # thus assumes data is missing at random (no pattern)
# m = 5 imputed datasets made for stabilized guessing

imputation <- mice(NHANES, m = 1, method = "pmm", seed = 123)
NHANES <- complete(imputation)

# Export to make sure it's never ruined!
NHANES_cleaned <- write.csv(NHANES, "NHANES_cleaned.csv")
```

Documentation:

-   [Albumin-Creatinine](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/ALB_CR_L.htm)

-   [Biochemistry Panel](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/BIOPRO_L.htm)

-   [Blood Pressure](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/BPXO_L.htm)

-   [Body Measures](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/BMX_L.htm)

-   [Complete Blood Count](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/CBC_L.htm)

-   [Cigarettes](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/SMQ_L.htm)

-   [Demographics](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DEMO_L.htm#SEQN)

-   [Diabetes](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DIQ_L.htm)

-   [Diet](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DR2TOT_L.htm#DR2TP182)

-   [Environmental Pollutants](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/PBCD_L.htm)

-   [Medical Conditions](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/MCQ_L.htm#MCQ160b)

-   [Mental Health](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DPQ_L.htm#DPQ070)

-   [Physical Activity](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/PAQ_L.htm)

## Step #2: Decision Tree to SVM Model

We only have 258/6055 (\~4.26%) CHF cases, thus we need some way to navigate the noise. This is roughly better than the estimated CHF incidence across American adults (6.7 million / 267 million = 2.51%) in terms of Signal-to-Noise Ratio (SNR). As a result, we are invested in experimenting methods to reduce SNR via a rule-out model to eliminate sure-cases without noise ([*Decision Tree*]{.underline}) with **low-threshold, max-sensitivity** capturing as much true positives, then a rule-in model to select the signal ([*Support Vector Model*]{.underline}) with **high-threshold, max-specificity** to select as much true positives.

### Step #2.1: Get Decision Tree

Intuitively, we care more about a false negative than a false positive. That's because we're okay shoving the false positive into the second round, but [we need the false negative]{.underline} pass into the sound round. So I want to give extra penalty to misclassifying a false negative where I take a confusion matrix and applying penalty, before I pass another decision tree, where heuristically, a false negative is 10x worse than a false positive:

|                     | **Predicted No CHF**    | **Predicted Yes CHF**  |
|---------------------|-------------------------|------------------------|
| **True No** **CHF** | 0                       | 1 \[*False Positive*\] |
| **True Yes CHF**    | 10 \[*False Negative*\] | 0                      |

Apparently, this is already well-known (whatever Claude), termed a **LOSS MATRIX**, part of rpart.

```{r}

NHANES_cleaned <- read.csv("NHANES_cleaned.csv")

# Drop 1st column (it's just row number) and factorize CHF
NHANES_cleaned <- NHANES_cleaned[, -1] %>%
  mutate(CHF = factor(CHF, levels = c("0", "1"), labels = c("No", "Yes")))

# Do 80/20 split while preserving CHF prevalence
set.seed(123)
split  <- initial_split(NHANES_cleaned, prop = 0.80, strata = CHF)
train  <- training(split)
test   <- testing(split)

###############################
# See loss matrix commentary above--we make the penalty matrix
loss_matrix <- matrix(c(0, 1, 10, 0), nrow = 2, byrow = TRUE)

###############################
# Initialize tree, fix hyperparameters to get idea
# Why those hyperparameters? I'm ok getting as many false positives (nonimportant    # predictors involved in the decision tree) to see obvious true positives.
  # Hence, I'm ok if the tree is (1) deep, (2) easy to split, and (3) complex
tree <- decision_tree(
  tree_depth  = 10,   # shallow (how many splits allowed)
  min_n       = 20,   # minimum obs to split node (like speciation)
  cost_complexity = 0.001) %>% # see notes; allows complexitiy, Gini Cross-Entropy
  set_engine("rpart", parms = list(loss = loss_matrix)) %>%
  set_mode("classification")   # Predictor is categorical

###############################
# Get the working df with the tree
working_df <- workflow() %>%
  add_formula(CHF ~ .) %>%
  add_model(tree)

fitted_df <- fit(working_df, data = train)

###############################
# Make it crazy big and read the png
  # So each node as in order of 1, 2, 3, 4 top-to-bottom:
  # 1: Classify Y/N, 2: P(No), 3: P(Yes), 4: % of training sample
  # Can't read tree, so put into png to be zoominable from there
png("tree_plot.png", width = 3000, height = 2000, res = 150)
rpart.plot(
  fitted_df$fit$fit$fit,
  type = 4,
  extra = 104,
  box.palette = "RdYlGn",
  main = "Stage 1 Decision Tree (Rule-Out)"
)
dev.off()

###############################
# Read the output directly
  # 1st entry: Condition, 2nd: Total obs...
  # 3rd: Loss if I predicted majority class here (hence all multiples of 10)
  # * denote terminal node (leaf = vertex degree 1)
  # So based off png and model, 87% of data has...
  # MCQ160C, E, and P >= 1.5, where 4222 No, 55 Yes (recall 0.8*258 = 206 Yes)
fitted_df$fit$fit$fit

# See what features I should care about
  # Importance = sum(abs(reductions in Gini Entropy per split))
  # https://www.geeksforgeeks.org/machine-learning/
      # gini-impurity-and-entropy-in-decision-tree-ml/
  # MCQ160C: Coronary Heart Disease
  # MCQ160E: Heart Attack
  # MCQ160P: COPD, Emphysema, Chronic Bronchitis
  # RIDAGEYR: Age of participants
  # LBXRDW:  Red blood cell distribution width
  # MCQ160D: Angina Pectoris (weaker version of Heart Attack)
png("tree_VIP.png", width = 800, height = 400, res = 120)
vip(fitted_df, num_features = 20, geom = "col") +
  labs(title = "Top 20 Features by Importance for Decision Tree") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
dev.off()
```

### Step #2.2: Be More Rigorous with Hyperparameters

Can't get it to work ;(

```{r}

# tree <- decision_tree(
#   tree_depth      = tune(),
#   min_n           = tune(),
#   cost_complexity = tune()) %>%
#   set_engine("rpart", parms = list(loss = loss_matrix)) %>%
#   set_mode("classification")
# 
# working_tune <- workflow() %>%
#   add_formula(CHF ~ .) %>%
#   add_model(tree)
# 
# # 5-fold CV (not 10 to ease computation), do stratified to maintain CHF prevalence
# # 64 * 5 = 320 models X__X RIP my laptop
# folds <- vfold_cv(train, v = 5, strata = CHF)
# 
# # Data-driven as (10, 20, 0.001) got me good stuff, so look around
# grid <- grid_regular(
#   tree_depth(range = c(5, 15)),
#   min_n(range = c(10, 50)),
#   cost_complexity(range = c(-4, -2)), # is log10, so 0.0001 to 0.01
#   levels = 4                          # 4 per parameters -> 4^3 = 64 combos
# )
# 
# # Want threshold-independent measure, w/ high sens + spec, so use ROC AUC
# tune_results <- tune_grid(
#   working_tune,
#   resamples = folds,
#   grid = grid,
#   metrics = metric_set(roc_auc),
#   control = control_grid(event_level = "second")
# )
# 
# autoplot(tune_results)
# # We get high ROC AUC for (I) more nodes, (II) less depth, (III) higher cost
# 
# best_tree_params <- tune_results %>% select_best(metric = "roc_auc")
# best_tree_params # Reflects autoplot inference
#
# # Get the tree from the chosen hyperparameters
#   # finalize_workflow: Fits tree given chosen hyperparameters
# final_workflow <- working_tune %>% finalize_workflow(best_tree_params)
# fitted_df <- fit(final_workflow, data = train)
```

### Step #2.3: Obtain and Apply the Tree-Based Model

```{r}

# 3 column df: .pred_No, .pred_Yes, and ground truth CHF per participant
predict_df <- predict(fitted_df, train, type = "prob") %>%
  bind_cols(select(train, CHF))

threshold_table <- predict_df %>% roc_curve(truth = CHF, .pred_Yes, 
                                            event_level = "second")

rule_out_threshold <- threshold_table %>%
  filter(sensitivity >= 0.90) %>%   # Get at least 90% of true CHF (sensitivity)
  slice_max(specificity, n = 1) %>% # Max true negatives ruled out (specificity) 
  pull(.threshold)                  # Get that threshold

rule_out_threshold

###############################
# Plot what exactly is threshold_table
png("rule_out_tree.png", width = 800, height = 400, res = 120)
threshold_table %>%
  pivot_longer(cols = c(sensitivity, specificity),
               names_to = "metric", values_to = "value") %>%
  ggplot(aes(x = .threshold, y = value, colour = metric)) +
  geom_line(linewidth = 2) +
  geom_vline(xintercept = rule_out_threshold,
             linetype = "dashed", colour = "red", linewidth = 2) +
  labs(title = "Sensitivity & Specificity by Rule-Out Threshold",
       x = "Threshold", y = NULL, colour = NULL) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
dev.off()

###############################
# Apply threshold to rule-out
train_another <- predict_df %>%
  bind_cols(train %>% select(-CHF)) %>%
  mutate(Stage1 = ifelse(.pred_Yes >= rule_out_threshold, "Yes", "No")) %>%
  filter(Stage1 == "Yes") %>%
  select(-starts_with(".pred"), -Stage1)

# Very good! Most true CHF retained (absolute) with increased SNR (relative)
train_another %>% count(CHF, name = "Count")
mean(train_another$CHF == "Yes")

# Tried to get the rules for ruling-out but that's for another time...
```

### Step #2.4: Get Support-Vector Machine

```{r}

# Preprocess columns
recipe_svm <- recipe(CHF ~ ., data = train_another) %>%
  step_zv(all_predictors()) %>%            # Remove 0 variance columns
  step_normalize(all_numeric_predictors()) # Scale data for SVM

# Radial Basis Function (RBF) Kernel
svm_specifity <- svm_rbf(
  cost      = tune(),        # Regularisation
  rbf_sigma = tune()) %>%    # Kernel bandwidth
  set_engine("kernlab") %>%  # Kernel-based SVM
  set_mode("classification")

working_svm_df <- workflow() %>%
  add_recipe(recipe_svm) %>%
  add_model(svm_specifity)

###############################
# 5-fold CV, stratified on CHF
folds <- vfold_cv(train_another, v = 5, strata = CHF)

# Random search over a small space
svm_grid <- grid_random(
  cost(range = c(-1, 2)),          # 10^(-1) to 10^2
  rbf_sigma(range = c(-3, -1)),    # 10^(-3) to 10^(-1)
  size = 15
)

# Tune with sensitivity
svm_results <- tune_grid(
  working_svm_df,
  resamples = folds,
  grid      = svm_grid,
  metrics   = metric_set(roc_auc, sensitivity, precision),
  control   = control_grid(save_pred = TRUE)
)

###############################
autoplot(svm_results)

# Best by ROC AUC
best_svm <- select_best(svm_results, metric = "roc_auc")
best_svm

final_svm     <- finalize_workflow(working_svm_df, best_svm)
fitted_svm_df <- fit(final_svm, data = train_another)
```

### Step #2.5: Obtain and Apply SVM-Based Model

```{r}

# Basically same code from #2.3 adapted to SVM
predict_svm_df <- predict(fitted_svm_df, train_another, type = "prob") %>%
  bind_cols(select(train_another, CHF))

threshold_svm <- predict_svm_df %>%
  roc_curve(truth = CHF, .pred_Yes, event_level = "second")

rule_in_threshold <- threshold_svm %>%
  # Lax sensitivty >= 0.7, but highest specificity for precision
  filter(sensitivity >= 0.7) %>%
  slice_max(specificity, n = 1) %>%
  pull(.threshold)

rule_in_threshold

###############################
# Basically same plot as 1st model
png("rule_in_SVM.png", width = 800, height = 400, res = 120)
threshold_svm %>%
  select(.threshold, sensitivity, specificity) %>%
  pivot_longer(cols = c(sensitivity, specificity),
               names_to = "metric", values_to = "value") %>%
  ggplot(aes(x = .threshold, y = value, colour = metric)) +
  geom_line(linewidth = 2) +
  geom_vline(xintercept = rule_in_threshold,
             linetype = "dashed", colour = "red", linewidth = 2) +
  labs(title = "Sensitivity & Specificity by Rule-In Threshold",
       x = "Threshold", y = NULL, colour = NULL) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
dev.off()

###############################
# Apply threshold to rule-in
final_df <- predict_svm_df %>%
  bind_cols(train_another %>% select(-CHF)) %>%
  mutate(Stage2 = ifelse(.pred_Yes >= rule_in_threshold, "Yes", "No")) %>%
  filter(Stage2 == "Yes") %>%
  select(-starts_with(".pred"), -Stage2)

# Alright, most CHF retained with only a few false positives, some false negatives
final_df %>% count(CHF, name = "Count")
mean(final_df$CHF == "Yes")
```

### Step #2.6: Evaluate on Test Set

```{r}

###############################
# 1) Decision tree to test set
###############################
test_tree_df <- predict(fitted_df, test, type = "prob") %>%
  bind_cols(select(test, CHF))

test_flagged_idx <- which(test_tree_df$.pred_Yes >= rule_out_threshold)
test_flagged     <- test[test_flagged_idx, ]

test %>% count(CHF, name = "Count")
test_flagged %>% count(CHF, name = "Count")

###############################
# 2) SVM to test set
###############################
test_svm_df <- predict(fitted_svm_df, test_flagged, type = "prob") %>%
  bind_cols(select(test_flagged, CHF))

# Final prediction vector for the whole test set
final_pred  <- rep("No", nrow(test)) # Initialize
svm_call    <- if_else(test_svm_df$.pred_Yes >= rule_in_threshold, "Yes", "No")
final_pred[test_flagged_idx]  <- svm_call

final_pred  <- factor(final_pred, levels = c("No", "Yes"))
truth       <- test$CHF

###############################
# 3) Moment of truth...
###############################
results <- tibble(truth = truth, estimate = final_pred)
sens  <- sensitivity(results,  truth, estimate, event_level = "second")
spec  <- specificity(results,  truth, estimate, event_level = "second")
ppv   <- ppv(results,          truth, estimate, event_level = "second")
npv   <- npv(results,          truth, estimate, event_level = "second")
bacc  <- bal_accuracy(results, truth, estimate, event_level = "second")

bind_rows(sens, spec, ppv, npv, bacc) %>%
  select(.metric, .estimate) %>% print()

# Confusion matrix is pretty bad...
conf_mat(results, truth, estimate) %>% autoplot(type = "heatmap") +
  labs(title = "Test Confusion Matrix for Decision Tree to SMV") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```
